Squintability and Other Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimization Choices

### 2025-11-23 revision/response.qmd

7. I was also wondering how the dependence of the smoothness on d can be explained, since it is not intuitive that this should change so much? (Note also that the last sentence in the caption for Table 1 seems to mix up squintability and smoothness.)

\textcolor{blue}{Table 1 is a subset of the estimated PP index, from the full table in the appendix. It happens that the three chosen ones all decrease as $n$ goes larger. This is not generally a pattern so this has been dropped from the caption.}

8. Finally, I appreciate the practical advice in section 7. Here, 7.1 should be somewhat extended in my opinion: does the planned tour then run through all the individual paths, should we follow the most successful jellyfish, or how would we pick?

\textcolor{blue}{In Section 7.1 some more details of using the planned tour with the jellyfish optimiser are given. More details on this are provided on the tourr package website, so that it can be kept updated if there are changes in the package.}

### 2025-11-23 jso.qmd
fig-alt AI generated added

```{r}
#| label: fig-example-functions
#| echo: false
#| fig-width: 9
#| fig-height: 6
#| out-width: 80%
#| fig-cap: "Examples of PP indexes with large (top row) and small (bottom row) squint angles, shown with a Huber plot, and histogram of the projected data corresponding to the optimal projection. A Huber plot shows the PP index values for all 1D data projections in polar coordinates. The skewness index on the trimodal data is also smoother than the binned normality index on RANDU."
#proj <- animate_xy(flea[,1:6], guided_tour(lda_pp(flea$species)))
best_proj <- matrix(c(-0.44157005,  0.80959831,
                      0.14622089, -0.08228193,
                      -0.08454314, -0.23234287,
                      0.70367852,  0.28185021,
                      0.27876825,  0.44819691,
                      0.45123453,  0.05896646), ncol=2, byrow=T)

>>>>

```{r}
#| label: fig-example-functions
#| echo: false
#| fig-width: 9
#| fig-height: 6
#| out-width: 80%
#| fig-cap: "Examples of PP indexes with large (top row) and small (bottom row) squint angles, shown with a Huber plot, and histogram of the projected data corresponding to the optimal projection. A Huber plot shows the PP index values for all 1D data projections in polar coordinates. The skewness index on the trimodal data is also smoother than the binned normality index on RANDU."
#| fig-alt: "The figure contains four panels. Panels (a) and (c) are polar coordinate plots showing the values of projection pursuit (PP) indexes—skewness for trimodal data in (a) and binned normality for RANDU data in (c)—across all 1D projections of 2D datasets. The radius represents the index value for each projection angle, with a dashed circle indicating the average index and a straight line marking the optimal projection. Panel (a) displays a smooth but non-convex skewness index with three pronounced modes and a wide squint angle, while panel (c) shows a noisier, non-normality index with less smoothness and a narrow squint angle. Panels (b) and (d) are histograms representing data projected onto the optimal directions identified in (a) and (c), respectively; panel (b) reveals trimodality and panel (d) shows the discrete nature of the RANDU data, which is harder to discern except near the optimal projection."
#proj <- animate_xy(flea[,1:6], guided_tour(lda_pp(flea$species)))
best_proj <- matrix(c(-0.44157005,  0.80959831,
                      0.14622089, -0.08228193,
                      -0.08454314, -0.23234287,
                      0.70367852,  0.28185021,
                      0.27876825,  0.44819691,
                      0.45123453,  0.05896646), ncol=2, byrow=T)

### 2025-11-23 jso.tex
fig-alt AI generated added

\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, rightrule=.15mm, colback=white, toptitle=1mm, opacityback=0, arc=.35mm, left=2mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, bottomrule=.15mm, coltitle=black, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, toprule=.15mm, breakable, colbacktitle=quarto-callout-note-color!10!white]

>>>>

\begin{tcolorbox}[enhanced jigsaw, titlerule=0mm, opacityback=0, left=2mm, colback=white, leftrule=.75mm, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, bottomtitle=1mm, breakable, rightrule=.15mm, arc=.35mm, toptitle=1mm, bottomrule=.15mm, opacitybacktitle=0.6, colframe=quarto-callout-note-color-frame, coltitle=black]
